---
title: "P8105 Homework 6: Linear Models"
author: "Annie Clark"
date: "November 16, 2018"
output: github_document
---


```{r}
library(tidyverse)
library(forcats)
library(purrr)
library(viridis)
library(glmnet)
library(modelr)

```


# Problem 1

The Washington Post has gathered data on homicides in 50 large U.S. cities and 
made the data available through a GitHub repository here. You can read their 
accompanying article here.

## Part 1: Load the data

```{r}
homicide_url = "https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"

## Read in data from github
homicide_data = 
  read.csv(url(homicide_url)) %>% 
  janitor::clean_names() %>% 
  as_tibble()
```

## Part 2: Create some variables and clean data
Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable 
indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; 
and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this 
is a data entry mistake. Modifiy victim_race to have categories white and 
non-white, with white as the reference category. Be sure that  victim_age is numeric.

```{r}
## create city_state var and filter out cities
homicide_data =
  homicide_data %>% 
  mutate(city_state = str_c(city, state, sep = ", ")) %>% 
  filter(!(city_state %in% c("Dallas, TX", "Phoenix, AZ", 
                            "Kansas City, MO", "Tulsa, AL")))

## additional tidying  
homicide_data = 
  homicide_data %>% 
  mutate(solved = as.numeric(disposition == "Closed by arrest"),
         victim_race_white = ifelse(victim_race == "White", "white", "nonwhite"),
         victim_race_white = fct_relevel(victim_race_white, ref = "white"),
         victim_age = as.numeric(victim_age))
```

## Part 3: Fit a logistic regression model for Baltimore, MD
For the city of Baltimore, MD, use the glm function to fit a logistic regression 
with resolved vs unresolved as the outcome and victim age, sex and race (as just 
defined) as predictors. Save the output of glm as an R object; apply the broom::tidy 
to this object; and obtain the estimate and confidence interval of the adjusted 
odds ratio for solving homicides comparing black victims to white victims keeping 
all other variables fixed.

```{r}
balt_solved_logistic = 
  homicide_data %>% 
  filter (city_state == "Baltimore, MD") %>% 
  glm(solved ~ victim_age + victim_sex + victim_race_white, data = ., family = binomial())

  balt_solved_logistic %>% 
    broom::tidy(conf.int = TRUE, exponentiate = TRUE) %>% 
    janitor::clean_names() %>%
    mutate(term = recode(term, victim_sexMale = "victim_sex: male", 
                         victim_race_whitenonwhite = "victim_race: nonwhite")) %>% 
    select(Term = term, "Adj Odds Ratio" = estimate, "Adj Conf: Lower bound" = conf_low, 
           "Adj Conf: Upper bound" = conf_high) %>% 
    knitr::kable(digits = 3)
```

## Part 4: Fit a logistical regression model for each city
Now run glm for each of the cities in your dataset, and extract the adjusted odds 
ratio (and CI) for solving homicides comparing black victims to white victims. Do 
this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest 
as necessary to create a dataframe with estimated ORs and CIs for each city.

```{r}
solved_logistic_function = function(df) {
  glm_fit = glm(solved ~ victim_age + victim_sex + victim_race_white, 
      data = df, family = binomial())
  
  glm_fit %>% 
    broom::tidy() %>% 
    mutate(odds_ratio = exp(estimate),
           conf_low = exp(estimate - (1.96 * std.error)), 
           conf_high = exp(estimate + (1.96 * std.error)))
}

## cannot get broom::confint_tidy or broom_tidy(conf.int = TRUE) to work within funciton, therefore had to calculate confidence intervals manually; could have also not used a function (put glm and broom::tidy in the map function), but with not being able to pull the confidence interval, it was clearer to me to put everything in a function

homocide_odds_city = 
  homicide_data %>% 
  group_by(city_state) %>%       
  nest() %>% 
  mutate(log_models = map(data, solved_logistic_function)) %>% 
  select(-data) %>% 
  unnest() %>% 
  filter(term == "victim_race_whitenonwhite") %>% 
  select(city_state, odds_ratio, conf_low, conf_high)
```


## Part 5: Make a plot of the models by city
Create a plot that shows the estimated ORs and CIs for each city. Organize cities 
according to estimated OR, and comment on the plot.

```{r}
homocide_odds_city %>% 
  mutate(city_state = fct_reorder(city_state, desc(odds_ratio))) %>%
  ggplot(aes(x = city_state, y = odds_ratio, color = city_state)) + 
    geom_point() +
    geom_errorbar(aes(ymin = conf_low, ymax = conf_high, width = 0.2)) + 
    theme(axis.text.x = element_text(angle = 60, hjust = 1),
          legend.position = "none") + 
    labs(
        title = "Adjusted Odds Ratios of Non-White to Whites for Solved Homicides in 50 Large Cities",
        x = "City, State",
        y = "Adjusted Odds Ratio (Non-Whites to Whites)", 
        caption = "Data from the Washington Post"
      ) +
    viridis::scale_color_viridis(
      discrete = TRUE
    )
```

This plot shows the adjusted odds ratio of non-whites to whites for solved homicides with 95% confidence bands. An odds ratio of 1 means that there is equal risk of having solved homicides for both whites and nonwhites. An odds ratio of less than 1 means higher risk of having solved homicides for white than nonwhites and an odds ratio of greater than 1 means higher risk of having solved homicides for nonwhites than whites. 

The cities with the highest odds ratios (Tampa, Durham, Birmingham and San Bernadino) also have large confidence bands. This is consistent with these cities have a low number of solved and unsolved incidences (small observations could lead to higher variance). Of the 50 cities, there are `r homocide_odds_city %>% filter(conf_high >= 1) %>% nrow()` cities that have confidence intervals below 1, meaning that there is evidence that nonwhites have a lower proportion of unsolved homicides than whites (with 95% confidence). The cities with the lowest odds ratios also have very narrow confidence bands, meaning that there is great certainty that in these cities (Boston, Ohmaha, Oakland, Pitssburgh), there is great evidence that cases of nonwhite victims go more unsolved than cases with white victims. 


# Problem 2
In this probelm, you will analyze data gathered to understand the effects of several 
variables on a child’s birthweight. This dataset, available here, consists of 
roughly 4000 children and includes the following variables:

* babysex: baby’s sex (male = 1, female = 2)
* bhead: baby’s head circumference at birth (centimeters)
* blength: baby’s length at birth (centimeteres)
* bwt: baby’s birth weight (grams)
* delwt: mother’s weight at delivery (pounds)
* fincome: family monthly income (in hundreds, rounded)
* frace: father’s race (1= White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other, 9 = Unknown)
* gaweeks: gestational age in weeks
* malform: presence of malformations that could affect weight (0 = absent, 1 = present)
* menarche: mother’s age at menarche (years)
* mheigth: mother’s height (inches)
* momage: mother’s age at delivery (years)
* mrace: mother’s race (1= White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other)
* parity: number of live births prior to this pregnancy
* pnumlbw: previous number of low birth weight babies
* pnumgsa: number of prior small for gestational age babies
* ppbmi: mother’s pre-pregnancy BMI
* ppwt: mother’s pre-pregnancy weight (pounds)
* smoken: average number of cigarettes smoked per day during pregnancy
* wtgain: mother’s weight gain during pregnancy (pounds)

## Part 1: Load the data

Load and clean the data for regression analysis (i.e. convert numeric to factor 
where appropriate, check for missing data, etc.).

```{r}
birthweight_data = read_csv(file = "./data/birthweight.csv") %>% 
  janitor::clean_names() %>% 
  mutate(babysex = recode(babysex, `1` = "male", `2` = "female"), 
         babysex = as.factor(babysex), 
         frace = recode(frace,  `1` = "white", `2` = "black", `3` = "asian", 
                        `4` = "puerto_rican", `8` = "other", `9` = "unknown"),
         frace = as.factor(frace), 
         mrace = recode(mrace,  `1` = "white", `2` = "black", `3` = "asian", 
                        `4` = "puerto_rican", `8` = "other" ),
         mrace = as.factor(mrace), 
         malform = as.factor(malform)) %>% 
  select(bwt, everything())

##birthweight_data %>% 
##  skimr::skim()

## not evaluating this because it does not render well in github; this shows that there are no missing values in any of the variables
```

## Part 2: Fit a regression model predicting birthweight

Propose a regression model for birthweight. This model may be based on a hypothesized 
structure for the factors that underly birthweight, on a data-driven model-building 
process, or a combination of the two. Describe your modeling process and show a 
plot of model residuals against fitted values – use add_predictions and add_residuals 
in making this plot.

```{r}
##based on hypothesis, not data driven

btw_model_hyp = function(df) {
  lm(bwt ~ bhead + blength + delwt + gaweeks, data = df)
}

birthweight_model = btw_model_hyp(birthweight_data)

  
birthweight_data %>% 
  add_predictions(birthweight_model) %>% 
  add_residuals (birthweight_model) %>% 
  ggplot(aes(x = pred, y = resid)) + 
    geom_point(alpha = 0.3) +
  labs(
        title = "Hypothesized Regression Model: Residuals vs. Predicted Values",
        x = "Predicted Birthweight (g)",
        y = "Residuals"
      )
```

I chose to build my model as though I was the investigator, not necessarily the data scientist. I hypothesized a model based on the variables that I thought would be the best predictors of birthweight. This model includes baby head cirumference, baby length, mother's weight at delivery, and gestational age (weeks). Because this was purely hypothesize driven, no techniques were used for variable selection, etc. 

When graphing the residuals vs. predicted, we see a heavy cluster in the lower right corner on the graph, with a tail fanning out to the top left. This is not necessarily what we want to see in this plot (which is a random scatter). After seeing this plot, I would be interested in possibly doing a transformation, or try other variables in the model. 

## Part 3: Create additional models and compare using cross validation

Compare your model to two others:

* One using length at birth and gestational age as predictors (main effects only)
* One using head circumference, length, sex, and all interactions (including the 
three-way interaction) between these

Make this comparison in terms of the cross-validated prediction error; use 
crossv_mc and functions in purrr as appropriate.

```{r}
## create additional models

btw_model_sml = function(df) {
  lm(bwt ~ blength + gaweeks, data = df)
}

btw_model_int = function(df) {
  lm(bwt ~ bhead + blength + babysex +
       (bhead * blength) + (bhead * babysex) + (blength * babysex) +
       (bhead * blength * babysex), data = df)
}

```


```{r}
## cross validation

## split data and run three models on each split

set.seed(10)

cv_birthweight = 
  crossv_mc(birthweight_data, 100) 

cv_birthweight = 
  cv_birthweight %>% 
  mutate(btw_model_hyp = map(train, btw_model_hyp),
         btw_model_sml = map(train, btw_model_sml),
         btw_model_int = map(train, btw_model_int)) %>% 
  mutate(rmse_btw_model_hyp = map2_dbl(btw_model_hyp, test, ~rmse(model = .x, data = .y)),
         rmse_btw_model_sml = map2_dbl(btw_model_sml, test, ~rmse(model = .x, data = .y)),
         rmse_btw_model_int = map2_dbl(btw_model_int, test, ~rmse(model = .x, data = .y)))


## visualize RMSEs by model
cv_birthweight %>% 
  select(starts_with("rmse")) %>% 
  gather(key = model, value = rmse) %>% 
  mutate(model = str_replace(model, "rmse_", ""),
         model = recode(model, btw_model_hyp = "Hypothesized", btw_model_sml = "Small",
                        btw_model_int = "Interaction"), 
         model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse, color = model)) + 
  geom_violin() +
  theme(legend.position = "none") +
  labs(
        title = "Comparing predictive birthweight models by RMSE",
        x = "Model",
        y = "RMSE"
      ) +
    viridis::scale_color_viridis(
      discrete = TRUE
    )
```

When comparing the hypothesized, small and interaction models, the hypothesized model is the clear winner because it has the smallest RMSE. The small model has the largest RMSEs and the interaction model has RMSEs similar (but not as low) to the hypothesized model. But because the interaction model is so much harder to interpret, the hypothesized model is a much better choice to use. 