---
title: "P8105 Homework 6: Linear Models"
author: "Annie Clark"
date: "November 16, 2018"
output: github_document
---
Things to work on:

* comments
* question 1: figure out how to get confidence interval to work 
* question 2: figure out lasso for building model


```{r}
library(tidyverse)
library(forcats)
library(purrr)
library(viridis)
library(glmnet)
library(modelr)

```


# Problem 1

The Washington Post has gathered data on homicides in 50 large U.S. cities and 
made the data available through a GitHub repository here. You can read their 
accompanying article here.

## Part 1: Load the data

```{r}
homicide_url = "https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"

## Read in data from github
homicide_data = 
  read.csv(url(homicide_url)) %>% 
  janitor::clean_names() %>% 
  as_tibble()
```

## Part 2: Create some variables and clean data
Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable 
indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; 
and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this 
is a data entry mistake. Modifiy victim_race to have categories white and 
non-white, with white as the reference category. Be sure that  victim_age is numeric.

```{r}
## create city_state var and filter out cities
homicide_data =
  homicide_data %>% 
  mutate(city_state = str_c(city, state, sep = ", ")) %>% 
  filter(!(city_state %in% c("Dallas, TX", "Phoenix, AZ", 
                            "Kansas City, MO", "Tulsa, AL")))

## additional tidying  
homicide_data = 
  homicide_data %>% 
  mutate(solved = as.numeric(disposition == "Closed by arrest"),
         victim_race_white = ifelse(victim_race == "White", "white", "nonwhite"),
         victim_race_white = fct_relevel(victim_race_white, ref = "white"),
         victim_age = as.numeric(victim_age))
```

## Part 3: Fit a logistic regression model for Baltimore, MD
For the city of Baltimore, MD, use the glm function to fit a logistic regression 
with resolved vs unresolved as the outcome and victim age, sex and race (as just 
defined) as predictors. Save the output of glm as an R object; apply the broom::tidy 
to this object; and obtain the estimate and confidence interval of the adjusted 
odds ratio for solving homicides comparing black victims to white victims keeping 
all other variables fixed.

```{r}
balt_solved_logistic = 
  homicide_data %>% 
  filter (city_state == "Baltimore, MD") %>% 
  glm(solved ~ victim_age + victim_sex + victim_race_white, data = ., family = binomial())

  balt_solved_logistic %>% 
    broom::tidy(conf.int = TRUE, exponentiate = TRUE) %>% 
    janitor::clean_names() %>%
    mutate(term = recode(term, victim_sexMale = "victim_sex: male", 
                         victim_race_whitenonwhite = "victim_race: nonwhite")) %>% 
    select(Term = term, "Adj Odds Ratio" = estimate, "Adj Conf: Lower bound" = conf_low, 
           "Adj Conf: Upper bound" = conf_high) %>% 
    knitr::kable(digits = 3)

```

## Part 4: Fit a logistical regression model for each city
Now run glm for each of the cities in your dataset, and extract the adjusted odds 
ratio (and CI) for solving homicides comparing black victims to white victims. Do 
this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest 
as necessary to create a dataframe with estimated ORs and CIs for each city.

```{r}
solved_logistic_function = function(df) {
  glm_fit = glm(solved ~ victim_age + victim_sex + victim_race_white, 
      data = df, family = binomial())
  
  glm_fit %>% 
    broom::tidy() %>% 
    mutate(odds_ratio = exp(estimate),
           conf_low = exp(estimate - (1.96 * std.error)), 
           conf_high = exp(estimate + (1.96 * std.error)))
}

## cannot get broom::confint_tidy or broom_tidy(conf.int = TRUE) to work within funciton, therefore had to calculate standard deviations by hand


homocide_odds_city_function = 
  homicide_data %>% 
  group_by(city_state) %>%       
  nest() %>% 
  mutate(log_models = map(data, solved_logistic_function)) %>% 
  select(-data) %>% 
  unnest() %>% 
  filter(term == "victim_race_whitenonwhite") %>% 
  select(city_state, odds_ratio, conf_low, conf_high)
```


```{r}
glm_homicide_function = function(df) {
  glm(solved ~ victim_age + victim_sex + victim_race_white, data = df, family = binomial())
}


homocide_odds_city = 
  homicide_data %>% 
  group_by(city_state) %>%       
  nest() %>% 
  mutate(log_models = map(.x = data, glm_homicide_function),
         tidy_models = map(log_models, broom::tidy))  %>% 
  select(-data, -log_models) %>% 
  unnest() %>% 
  janitor::clean_names() %>% 
  mutate(conf_low = estimate - (1.96 * std_error), 
         conf_high = estimate + (1.96 * std_error)) %>% 
  filter(term == "victim_race_whitenonwhite") %>% 
  select(city_state, estimate, conf_low, conf_high)


broom_function = function(df) {
  df %>% 
    broom::tidy(conf.int = TRUE, exponentiate = TRUE)
}


  homicide_data %>% 
  group_by(city_state) %>%       
  nest() %>% 
  mutate(log_models = map(.x = data, glm_homicide_function),
         tidy_models = map(log_models, broom_function))  %>% 
  select(-data, -log_models) %>% 
  unnest() %>% 
  janitor::clean_names() %>% 
  mutate(odds_ratio = exp(estimate), 
         conf_low = exp(conf_low), 
         conf_high = exp(conf_high)) %>% 
  filter(term == "victim_race_whitenonwhite") %>% 
  select(city_state, odds_ratio, conf_low, conf_high)
```

## Part 5: Make a plot of the models by city
Create a plot that shows the estimated ORs and CIs for each city. Organize cities 
according to estimated OR, and comment on the plot.

```{r}
homocide_odds_city %>% 
  mutate(city_state = fct_reorder(city_state, desc(estimate))) %>%
  ggplot(aes(x = city_state, y = estimate, color = city_state)) + 
    geom_point() +
    geom_errorbar(aes(ymin = conf_low, ymax = conf_high, width = 0.2)) + 
    theme(axis.text.x = element_text(angle = 60, hjust = 1),
          legend.position = "none") + 
    labs(
        title = "Adjusted Odds Ratios of Non-White to Whites for Solved Homicides in 50 Large Cities",
        x = "City, State",
        y = "Adjusted Odds Ratio (Non-Whites to Whites)", 
        caption = "Data from the Washington Post"
      ) +
    viridis::scale_color_viridis(
      discrete = TRUE
    )
```


```{r}
homocide_odds_city_function %>% 
  mutate(city_state = fct_reorder(city_state, desc(odds_ratio))) %>%
  ggplot(aes(x = city_state, y = odds_ratio, color = city_state)) + 
    geom_point() +
    geom_errorbar(aes(ymin = conf_low, ymax = conf_high, width = 0.2)) + 
    theme(axis.text.x = element_text(angle = 60, hjust = 1),
          legend.position = "none") + 
    labs(
        title = "Adjusted Odds Ratios of Non-White to Whites for Solved Homicides in 50 Large Cities",
        x = "City, State",
        y = "Adjusted Odds Ratio (Non-Whites to Whites)", 
        caption = "Data from the Washington Post"
      ) +
    viridis::scale_color_viridis(
      discrete = TRUE
    )
```

# Problem 2
In this probelm, you will analyze data gathered to understand the effects of several 
variables on a child’s birthweight. This dataset, available here, consists of 
roughly 4000 children and includes the following variables:

* babysex: baby’s sex (male = 1, female = 2)
* bhead: baby’s head circumference at birth (centimeters)
* blength: baby’s length at birth (centimeteres)
* bwt: baby’s birth weight (grams)
* delwt: mother’s weight at delivery (pounds)
* fincome: family monthly income (in hundreds, rounded)
* frace: father’s race (1= White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other, 9 = Unknown)
* gaweeks: gestational age in weeks
* malform: presence of malformations that could affect weight (0 = absent, 1 = present)
* menarche: mother’s age at menarche (years)
* mheigth: mother’s height (inches)
* momage: mother’s age at delivery (years)
* mrace: mother’s race (1= White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other)
* parity: number of live births prior to this pregnancy
* pnumlbw: previous number of low birth weight babies
* pnumgsa: number of prior small for gestational age babies
* ppbmi: mother’s pre-pregnancy BMI
* ppwt: mother’s pre-pregnancy weight (pounds)
* smoken: average number of cigarettes smoked per day during pregnancy
* wtgain: mother’s weight gain during pregnancy (pounds)

## Part 1: Load the data

Load and clean the data for regression analysis (i.e. convert numeric to factor 
where appropriate, check for missing data, etc.).

```{r}
birthweight_data = read_csv(file = "./data/birthweight.csv") %>% 
  janitor::clean_names() %>% 
  mutate(babysex = recode(babysex, `1` = "male", `2` = "female"), 
         babysex = as.factor(babysex), 
         frace = recode(frace,  `1` = "white", `2` = "black", `3` = "asian", 
                        `4` = "puerto_rican", `8` = "other", `9` = "unknown"),
         frace = as.factor(frace), 
         mrace = recode(mrace,  `1` = "white", `2` = "black", `3` = "asian", 
                        `4` = "puerto_rican", `8` = "other" ),
         mrace = as.factor(mrace), 
         malform = as.factor(malform)) %>% 
  select(bwt, everything())

skimr::skim(birthweight_data)
## no missing data
```

## Part 2: Fit a regression model predicting birthweight

Propose a regression model for birthweight. This model may be based on a hypothesized 
structure for the factors that underly birthweight, on a data-driven model-building 
process, or a combination of the two. Describe your modeling process and show a 
plot of model residuals against fitted values – use add_predictions and add_residuals 
in making this plot.

```{r}

##based on hypothesis, not data driven


birthweight_model_function = function(df) {
  lm(bwt ~ bhead + blength + delwt + gaweeks, data = df)
}

birthweight_model = birthweight_model_function(birthweight_data)

  
birthweight_data %>% 
  add_predictions(birthweight_model) %>% 
  add_residuals (birthweight_model) %>% 
  ggplot(aes(x = pred, y = resid)) + 
    geom_point(alpha = 0.3) +
  labs(
        title = "Hypothesized Regression Model: Residuals vs. Predicted Values",
        x = "Predicted Birthweight (g)",
        y = "Residuals"
      )
```


```{r, eval = FALSE}
##trying lasso

birthweight_lasso = glmnet(birthweight_data, birthweight_data$bwt, alpha = 1, lambda = lambda)

?glmnet

x = model.matrix(bwt ~ ., data = birthweight_data)[, -1]
y = birthweight_data$bwt
lasso.mod = glmnet(as.matrix(x), y, alpha = 1)

x <- model.matrix(Sepal.Length~., iris)[,-1]
y <- iris$Sepal.Length
mod <- cv.glmnet(as.matrix(x), y, alpha=1)

as.matrix(coef(lasso.mod, lasso.mod$lambda.min))


require(glmnet)

ezlasso=function(df,yvar,folds=10,trace=F,alpha=1){
  x<-model.matrix(as.formula(paste(yvar,"~.")),data=df)
  x=x[,-1] ##remove intercept

  glmnet1<-glmnet::cv.glmnet(x=x,y=df[,yvar],type.measure='mse',nfolds=folds,alpha=alpha)

  co<-coef(glmnet1,s = "lambda.1se")
  inds<-which(co!=0)
  variables<-row.names(co)[inds]
  variables<-variables[!(variables %in% '(Intercept)')];
  return( c(yvar,variables));
}

ezlasso(birthweight_data, birthweight_data$bwt)


ezlasso=function(df,yvar,folds=10,trace=F,alpha=1){
  x<-model.matrix(as.formula(paste(yvar,"~.")),data=df)
  x=x[,-1] ##remove intercept

  glmnet1<-glmnet::cv.glmnet(x=x,y=df[,yvar],type.measure='mse',nfolds=folds,alpha=alpha)

  co<-coef(glmnet1,s = "lambda.1se")
  inds<-which(co!=0)
  variables<-row.names(co)[inds]
  variables<-variables[!(variables %in% '(Intercept)')];
  return( c(yvar,variables));
}

x = model.matrix(bwt ~ ., data = birthweight_data)[, -1]
y = birthweight_data$bwt
lasso.mod = glmnet(as.matrix(x), y, alpha = 1)

plot(lasso.mod, xvar = "lambda")


coef(lasso.mod, s = 100)

crossval <-  cv.glmnet(x = X, y = y)
plot(crossval)
penalty <- crossval$lambda.min #optimal lambda
penalty #minimal shrinkage
fit1 <-glmnet(x = X, y = y, alpha = 1, lambda = penalty ) #estimate the model with that
coef(fit1)
```


## Part 3: Create additional models and compare using cross validation

Compare your model to two others:

* One using length at birth and gestational age as predictors (main effects only)
* One using head circumference, length, sex, and all interactions (including the 
three-way interaction) between these

```{r}
birthweight_model_alt1_function = function(df) {
  lm(bwt ~ blength + gaweeks, data = df)
}

## birthweight_model_alt1_function(birthweight_data)

birthweight_model_alt2_function = function(df) {
  lm(bwt ~ bhead + blength + babysex +
       (bhead * blength) + (bhead * babysex) + (blength * babysex) +
       (bhead * blength * babysex), data = df)
}

## birthweight_model_alt2_function(birthweight_data)

```

Make this comparison in terms of the cross-validated prediction error; use 
crossv_mc and functions in purrr as appropriate.

```{r}
cv_birthweight = 
  crossv_mc(birthweight_data, 100) 

cv_birthweight = 
  cv_birthweight %>% 
  mutate(birthweight_model = map(train, birthweight_model_function),
         birthweight_model_alt1 = map(train, birthweight_model_alt1_function),
         birthweight_model_alt2 = map(train, birthweight_model_alt2_function)) %>% 
  mutate(rmse_birthweight_model = map2_dbl(birthweight_model, test, ~rmse(model = .x, data = .y)),
         rmse_birthweight_model_alt1 = map2_dbl(birthweight_model_alt1, test, ~rmse(model = .x, data = .y)),
         rmse_birthweight_model_alt2 = map2_dbl(birthweight_model_alt2, test, ~rmse(model = .x, data = .y)))



cv_birthweight %>% 
  select(starts_with("rmse")) %>% 
  gather(key = model, value = rmse) %>% 
  mutate(model = str_replace(model, "rmse_", ""),
         model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse, color = model)) + 
  geom_violin() +
  theme(legend.position = "none") +
  labs(
        title = "Comparing predictive birthweight models by RMSE",
        x = "Model",
        y = "RMSE"
      ) +
    viridis::scale_color_viridis(
      discrete = TRUE
    )
```

